---
title: "Unique Features of R for Programmers"
author: "Yuhao, JC (mentor)"
mentor: "JC"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    incremental: false
---

## Introduction

- **R** offers powerful features that set it apart in the programming world.
- Let's explore some unique aspects that make R interesting for programmers.


## Vectorized Operations

- Perform operations on entire vectors without explicit loops.
- Leads to concise and efficient code.

```{r}
# Vectorized addition
a <- 1:5
b <- 6:10
result <- a + b
print(result)  # Outputs: 7 9 11 13 15
```



## Functional Programming Paradigms

- Functions are first-class citizens.
- Supports higher-order functions, closures, and function factories.


```{r}
# Function returning another function
power_factory <- function(n) {
  function(x) x ^ n
}
square <- power_factory(2)
cube <- power_factory(3)
square(4)  # Outputs: 16
cube(2)    # Outputs: 8
```



## Lazy Evaluation

- Arguments are evaluated only when needed.
- Enables flexible and efficient function definitions.


```{r}
# Lazy evaluation example
lazy_function <- function(x, y) {
  if (x > 0) x else y
}
result <- lazy_function(5, {print("y is evaluated"); 0})
print(result)  # Outputs: 5
# "y is evaluated" is not printed
```


## Metaprogramming Capabilities

- Modify and generate code at runtime.
- Use expressions and environments to manipulate code objects.


```{r}
# Creating a function from an expression
expr <- expression(function(x) x + 1)
increment <- eval(expr)
increment(5)  # Outputs: 6
```


## Non-Standard Evaluation (NSE)

- Functions can capture expressions instead of values.
- Simplifies syntax for data manipulation and modeling.

---

Let's first import dplyer:

```{r}
library(dplyr)
```

---

Then we can peform data manipulation simply using "verbs":

```{r}
# Using NSE in dplyr
result <- mtcars %>%
  select(mpg, cyl) %>%
  filter(cyl == 6)
print(result)
```


## Advanced Graphics with ggplot2 

- Create complex and customizable visualizations.
- Based on the **Grammar of Graphics**.

---

```{r}
library(ggplot2)

# Create the plot and store it in a variable
p <- ggplot(mtcars, aes(x = wt, y = mpg, color = factor(gear))) +
     geom_point(size = 3) +
     labs(
       title = "Fuel Efficiency by Car Weight",
       x = "Weight (1000 lbs)",
       y = "Miles per Gallon",
       color = "Gears"
     ) +
     theme_minimal()
```

---

```{r}
# Display the plot
p
```

## Interfacing with Other Languages

- Integrate C/C++ code for performance-critical sections using **Rcpp**.
- Interface with Python via **reticulate**.

---

```{r}
library(Rcpp)
cppFunction('
int fibonacci(int n) {
  if (n <= 1) return n;
  else return(fibonacci(n - 1) + fibonacci(n - 2));
}')
fibonacci(10)  # Outputs: 55
```


## Statistical Modeling and Machine Learning

- Extensive support for statistical analyses.
- Machine learning libraries like **caret**, **xgboost**, and **randomForest**.

---

```{r}
library(randomForest)
```

---
```{r}
model <- randomForest(Species ~ ., data = iris)
print(model)
```

## Parallel Computing

- Utilize multiple cores with packages like **parallel** and **foreach**.
- Simplifies parallelization of code.


```{r}
library(parallel)
results <- mclapply(1:5, function(x) x^2, mc.cores = 2)
print(results)  # Outputs: List of squared numbers
```


## Rich Ecosystem and Community

- Over 16,000 packages available on CRAN.
- Active community contributing to diverse domains.


```{r, eval=FALSE}
install.packages(c("dplyr", "ggplot2", "Rcpp", "randomForest", "parallel"))
```
